---
title: "Using Crosstalk"
date: "2023-01-23"
categories: [crosstalk, R, ggplot2, plotly, code]
execute: 
  warning: false
---

Most everyone who works with data is familiar with or at least aware of 
  [Shiny](https://shiny.rstudio.com/) and all of the great things you can do
  with it's interactive capabilities. However, Shiny also requires some extra work
  on the backend that may not be available to everyone. I can write Shiny apps
  all day, but I can't share them with coworkers since they don't have R/Rstudio,
  and my employer doesn't have a Shiny server, nor do we have an account with 
  [shinyapps.io](https://www.shinyapps.io/). 
  
Most of the time when I need to share some data that I have worked on in R I use
  Rmarkdown/Quarto, with a PDF or Word output for printing purposes and an HTML
  report that uses Plotly, DT, Knitr, etc. to give some level of responsiveness.
  However, the `crosstalk` library allows users to bridge the gap between
  a fully interactive Shiny document and a fixed Rmarkdown/Quarto document.


## Libraries

First things first, we'll need to load some libraries:

```{r}
#| label: libraries

library(tidyverse)
library(crosstalk)
library(d3scatter)
library(plotly)
library(DT)
library(summarywidget)

```

 - tidyverse: Needs no explanation
 - crosstalk: This is the magic sauce
 - d3scatter: Simple plots (think `ggplot() + theme_bw()`) that work with `crosstalk`
 - plotly: Wonderful interactive plots on their own, `plotly` plots can be made 
    more interactive with `crosstalk`
 - DT: DT makes tables that already have filtering and responsiveness built-in,
    but we can use crosstalk to further filter and update in sync with the plots
 - summarywidget: This lets us generate summary values, like you might do in Shiny


Other crosstalk compatible libraries are `leaflet` for mapping and `rgl` for 
  interactive 3d plotting.


## The Basics

To use crosstalk, we have to create a SharedData object:
```{r}
#| label: data-sharedMtCars
sharedMtCars <- SharedData$new(mtcars)
```

Then we use these objects like data.frames in crosstalk-compatible widgets.

`crosstalk` also provides some `filter_*` functions:
```{r}
#| label: basic-d3scatter-with-filters
carsCheckBox <- filter_checkbox("cyl", "Cylinders", sharedMtCars, ~cyl, inline=T)
carsSelect <- filter_select("am", "Automatic", sharedMtCars, ~am)
carsSlider <- filter_slider("disp", "Displacement", sharedMtCars, ~disp)

carsPlot <- d3scatter(sharedMtCars, ~wt, ~mpg, ~factor(cyl), width="100%")
```

Changing the selected values will cause this plot to update, no need for Shiny:

::: {.columns}

:::: {.column width="30%"}
```{r}
#| label: basic-filters
#| echo: false
carsCheckBox
carsSelect
carsSlider
```
::::

:::: {.column width="70%"}
```{r}
#| label: basic-plot
#| echo: false
carsPlot
```
::::

:::




## Simple Plots


### d3scatter

d3scatter is a dead-simple library for graphs that work with crosstalk. As seen
  above, it's very much like a simple ggplot in appearance. As long as the plot
  and all of the `filter_*` functions use the same SharedData object, then the
  plot will update in real time with the various selections.
  

What's really neat about crosstalk is that we can update multiple plots at once:

```{r}
#| label: simple-plots-filters-plots
# create a new SharedObject so we're not manipulating everything on the page:
sharedMtCars <- SharedData$new(mtcars)

carsCheckBox <- filter_checkbox("cyl", "Cylinders", sharedMtCars, ~cyl, inline=T)
carsSelect <- filter_select("am", "Automatic", sharedMtCars, ~am)
carsSlider <- filter_slider("disp", "Displacement", sharedMtCars, ~disp)

carsPlotMpg <- d3scatter(sharedMtCars, ~wt, ~mpg, ~factor(cyl), width="100%")
carsPlothp <- d3scatter(sharedMtCars, ~wt, ~hp, ~factor(cyl), width="100%")

```

::: {.columns}

:::: {.column width="30%"}
```{r}
#| label: simple-plots-filters
#| echo: false
carsCheckBox
carsSelect
carsSlider
```
::::

:::: {.column width="35%"}
```{r}
#| label: simple-plot-mpg
#| echo: false
carsPlotMpg
```
::::

:::: {.column width="35%"}
```{r}
#| label: simple-plot-hp
#| echo: false
carsPlothp
```
::::

:::

`d3scatter` also allows click-and-drag selection of points, which will also be 
  applied to any other `d3scatter` plots using the same SharedData.


### Plotly

Plotly graphs are already responsive and allow zooming and selection, but using 
  crosstalk `filter_` functions we can do even more selection than what would be
  available otherwise, with the effects carrying over into other plots. 

```{r}
#| label: plotly-filter-plots
sharedMtCars <- SharedData$new(mtcars)

carsCheckBox <- filter_checkbox("cyl", "Cylinders", sharedMtCars, ~cyl, inline=T)
carsSelect <- filter_select("am", "Automatic", sharedMtCars, ~am)
carsSlider <- filter_slider("disp", "Displacement", sharedMtCars, ~disp)

plotlyMpg <- plot_ly(sharedMtCars, x=~wt, y=~mpg, color = ~factor(cyl))
plotlyHp <- plot_ly(sharedMtCars, x=~wt, y=~hp, color = ~factor(cyl))
```

::: {.columns}

:::: {.column width="20%"}
```{r}
#| label: plotly-filters
#| echo: false
carsCheckBox
carsSelect
carsSlider
```
::::

:::: {.column width="40%"}
```{r}
#| label: plotly-plot-mpg
#| echo: false
plotlyMpg 
```
::::

:::: {.column width="40%"}
```{r}
#| label: plotly-plot-hp
#| echo: false
plotlyHp
```
::::

:::



