---
title: "Using Crosstalk"
date: "2023-01-23"
categories: [crosstalk, R, ggplot2, plotly, code]
execute: 
  warning: false
---

Most everyone who works with data is familiar with or at least aware of 
  [Shiny](https://shiny.rstudio.com/) and all of the great things you can do
  with it's interactive capabilities. However, Shiny also requires some extra work
  on the backend that may not be available to everyone. I can write Shiny apps
  all day, but I can't share them with coworkers since they don't have R/Rstudio,
  and my employer doesn't have a Shiny server, nor do we have an account with 
  [shinyapps.io](https://www.shinyapps.io/). 
  
Most of the time when I need to share some data that I have worked on in R I use
  Rmarkdown/Quarto, with a PDF or Word output for printing purposes and an HTML
  report that uses Plotly, DT, Knitr, etc. to give some level of quasi-interactivity.
  
However, the `crosstalk` library allows users to bridge the gap somewhat between
  a fully interactive Shiny document and a fixed Rmarkdown/Quarto document.


## The Basics

### Libraries

First things first, we'll need to load some libraries:

```{r}
#| label: libraries

library(tidyverse)
library(crosstalk)
library(d3scatter)
library(plotly)
library(rgl)
library(DT)
library(summarywidget)

```

 - Tidyverse: Needs no explanation
 - Crosstalk: This is the magic sauce
 - d3scatter: Simple plots (think `ggplot() + theme_bw()`) that work with `crosstalk`
 - plotly: Wonderful interactive plots on their own, `plotly` plots can be made 
    more interactive with `crosstalk`
 - rgl: 3D plots!
 - DT: DT makes tables that already have filtering and responsiveness built-in,
    but we can use crosstalk to further filter and update in sync with the plots
 - summarywidget: This lets us generate summary values, like you might do in Shiny


I don't use maps in my work, but leaflet can also use crosstalk:
```{r}
#| label: library-leaflet 
#| eval: false

library(leaflet)
```


### Simple Plots


```{r}
mtcars |> 
  ggplot(aes(x=wt, y=mpg, color=factor(cyl))) +
  geom_point() +
  geom_smooth(method="lm", se=F) +
  theme_bw()
```


```{r}
sharedMtCars <- SharedData$new(mtcars)


cylCheckbox <- filter_checkbox("cyl", "Cylinders", sharedMtCars, ~cyl, inline=T)
d3Plot <- d3scatter(sharedMtCars, ~wt, ~mpg, ~factor(cyl), width="100%")
```

::: {.columns}

:::: {.column width="20%"}
```{r}
#| echo: false
cylCheckbox
```
::::

:::: {.column width="80%"}
```{r}
#| echo: false
d3Plot
```
::::

:::



```{r}
sharedMtCars <- SharedData$new(mtcars)


cylCheckbox <- filter_checkbox("cyl", "Cylinders", sharedMtCars, ~cyl, inline=T)
plotlyPlot <- plot_ly(sharedMtCars, x=~wt, y=~mpg, color = ~factor(cyl))
```

::: {.columns}

:::: {.column width="30%"}
```{r}
#| echo: false
cylCheckbox
```
::::

:::: {.column width="70%"}
```{r}
#| echo: false
plotlyPlot
```
::::

:::


